---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(pryr)
library(dplyr)
library(zstdlite)
```

# zstdlite

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![R build status](https://github.com/coolbutuseless/zstdlite/workflows/R-CMD-check/badge.svg)](https://github.com/coolbutuseless/zstdlite/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

`zstdlite` provides access to the very fast (and highly configurable) 
 [zstd](https://github.com/facebook/zstd) library for performing in-memory compression of 
 numeric vectors and arrays.

The scope of this package is limited - it aims to provide functions for direct compression of vectors
and arrays
which contain raw, integer, real, complex or logical values.  It does this by operating on the
data payload within the vectors, and gains significant speed by not serializing the R object
itself.  If you wanted to 
compress arbitrary R objects, you must first manually convert into a raw
vector representation using `base::serialize()`.

As of v0.1.3, the dimensions of the input data is stored, and matrices and 
vectors are restored to their original shape upon decompression.


Currently zstd code provided with this package is v1.4.6.


### Design Choices

`zstdlite` will compress the *data payload* within a numeric-ish vector, and not the
R object itself.  Along with this payload, a minimum of meta information is kept
to store data type and array dimensions.

#### Limitations

* Any attribute information (other than array dimensions) is lost.
* Values must be of type: raw, integer, real, complex or logical.


## Installation

You can install from [GitHub](https://github.com/coolbutuseless/zstdlite) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/zstdlite)
```

## Basic usage of zstdlite

```{r}
arr <- array(1:27, c(3, 3, 3))
pryr::object_size(arr)


buf <- zstd_compress(arr)
length(buf) # Number of bytes

zstd_decompress(buf)
```





## Compressing 1 million Integers

`zstdlite` supports the direct compression of raw, integer, real, complex and logical vectors.

These vectors do not need to be serialized first to a raw representation, instead
the data-payload within these vectors is compressed.

```{r}
library(zstdlite)
library(lz4lite)

N                 <- 1e6
input_ints        <- sample(1:5, N, prob = (1:5)^2, replace = TRUE)
compressed_lo     <- zstd_compress(input_ints)
compressed_hi     <- zstd_compress(input_ints, level = 100)
compressed_lo_lz4 <- lz4_compress(input_ints, acc = 1)
compressed_hi_lz4 <- lz4_compress(input_ints, use_hc = TRUE, hc_level = 12)
```


```{r echo = FALSE}
lens <- c(
  length(zstd_compress(input_ints, level =  1)),
  length(zstd_compress(input_ints, level =  3)),
  length(zstd_compress(input_ints, level = 10)),
  length(zstd_compress(input_ints, level = 22)),
  length(compressed_lo_lz4),
  length(compressed_hi_lz4)
)
```



<details>
<summary> Click here to show/hide benchmark code </summary>

```{r}
library(zstdlite)

res <- bench::mark(
  zstd_compress(input_ints, level =   1),
  zstd_compress(input_ints, level =   3),
  zstd_compress(input_ints, level =  10),
  zstd_compress(input_ints, level =  22),
  lz4_compress (input_ints, acc = 1),
  lz4_compress (input_ints, use_hc = TRUE, hc_level = 12),
  check = FALSE
)
```
</details>


```{r echo = FALSE}
res %>% 
  mutate(`MB/s` = round(N*4/1024^2 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  mutate(compression_ratio = round(lens/(N*4), 3)) %>%
  select(expression, median, `itr/sec`, `MB/s`, compression_ratio) %>%
  mutate(package = c(
    rep('zstdlite', 4),
    rep('lz4lite', 2)
  )) %>%
  select(package, everything()) %>%
  knitr::kable()
```


### Decompressing 1 million integers


<details>
<summary> Click here to show/hide benchmark code </summary>

```{r}
res <- bench::mark(
  zstd_decompress(compressed_lo),
  zstd_decompress(compressed_hi),
  lz4_decompress(compressed_lo_lz4),
  lz4_decompress(compressed_hi_lz4),
  check = FALSE
)
```

</details>

```{r echo = FALSE}
res %>% 
  mutate(`MB/s` = round(N*4/1024^2 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  select(expression, median, `itr/sec`, `MB/s`)  %>%
  mutate(package = c(
    rep('zstdlite', 2),
    rep('lz4lite', 2)
  )) %>%
  select(package, everything()) %>%
  knitr::kable()
```



```{r echo = FALSE, eval = FALSE}
set.seed(1)
dat <- sample(1:10, 1e6, replace = TRUE)

res <- bench::press(
  level = seq(-10, 22),
  bench::mark(zstd_compress(dat, level = level))
)


tibble:::print.tbl(res, n = Inf)

```




## Technical bits

### Why only vectors of raw, integer, real, complex or logical?

R objects can be considered to consist of: 

* a header - giving information like length and information for the garbage collector
* a body - data of some kind.

The vectors supported by `zstdlite` are those vectors whose body consists of 
data that is directly interpretable as a contiguous sequence of bytes representing 
numerical values.

Other R objects (like lists or character vectors) are really collections of pointers 
to other objects, and do not live in memory as a contiguous sequence of byte data.

### How it works.

##### Compression

1. Given a pointer to a standard numeric vector from R (i.e. an *SEXP* pointer).
2. Ignore any attributes (except for dimensions) and compress the data payload within 
   the object.
3. Prefix the compressed data with 
    * a 4 byte header giving the SEXP type
    * the dimensions of the data
4. Return a raw vector to the user containing the compressed bytes.

##### Decompression

1. Strip off the 4-bytes of header information and the dimension information
2. Feed the other bytes in to the ZSTD decompression function written in C
3. Use the header to decide what sort of R object this is.
4. Decompress the data into an R object of the correct type, and set the
   dimensions 
5. Return the R object to the user.


### Framing of the compressed data

* `zstdlite` prefixes the standard Zstandard frame with some extra bytes.
* The compressed representation is the compressed data prefixed with 
  a custom 8 byte header consisting of 
    * 'ZST'
    * 1-byte for SEXP type i.e. INTSXP, RAWSXP, REALSXP or LGLSXP
* This data representation 
    * is compatible with the standard Zstandard frame 
      format if the leading bytes are removed.
    * is likely to evolve (so currently do not plan on compressing something in 
      one version of `zstdlite` and decompressing in another version.)


### Zstd "Single File" Libary

* To simplify the code within this package, it uses the 'single file library' version of zstd
* To update this package when zstd is updated, create the single file library version
    1. cd zstd/contrib/single_file_libs
    2. sh create_single_file_library.sh
    3. Wait.....
    4. copy zstd/contrib/single_file_libs/zstd.c into zstdlite/src
    5. copy zstd/lib/zstd.h into zstdlite/src

## Related Software


For a more general solution to fast serialization of R objects, see the 
[fst](https://github.com/fstpackage/fst) or [qs](https://cran.r-project.org/package=qs) packages.

* [lz4](https://github.com/lz4/lz4) and [zstd](https://github.com/facebook/zstd) - both by Yann Collet
* [fst](https://github.com/fstpackage/fst) for serialisation of data.frames using
  lz4 and zstd
* [qs](https://cran.r-project.org/package=qs) for fast serialization of arbitrary R objects
  with lz4 and zstd

## Acknowledgements

* Yann Collett for releasing, maintaining and advancing 
[lz4](https://github.com/lz4/lz4) and [zstd](https://github.com/facebook/zstd)
* R Core for developing and maintaining such a wonderful language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
