---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(dplyr)
library(zstdlite)
```

# zstdlite

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![R build status](https://github.com/coolbutuseless/zstdlite/workflows/R-CMD-check/badge.svg)](https://github.com/coolbutuseless/zstdlite/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R-CMD-check](https://github.com/coolbutuseless/zstdlite/workflows/R-CMD-check/badge.svg)](https://github.com/coolbutuseless/zstdlite/actions)
<!-- badges: end -->

`zstdlite` provides access to the very fast (and highly configurable) 
 [zstd](https://github.com/facebook/zstd) library for performing in-memory compression of 
R objects.

For rock solid general solutions to fast serialization of R objects, see the 
[fst](https://github.com/fstpackage/fst) or [qs](https://cran.r-project.org/package=qs) packages.

[zstd](https://github.com/facebook/zstd) code provided with this package is v1.5.0.

## What's in the box

* `zstd_serialize()` and `zstd_unserialize()` for converting R objects to/from
  a compressed representation

## Installation

You can install from [GitHub](https://github.com/coolbutuseless/zstdlite) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/zstdlite')
```

## Basic usage of zstdlite

```{r}
arr <- array(1:27, c(3, 3, 3))
lobstr::obj_size(arr)


buf <- zstd_serialize(arr)
length(buf) # Number of bytes

# compression ratio
length(buf)/as.numeric(lobstr::obj_size(arr))

zstd_unserialize(buf)
```





## Compressing 1 million Integers


```{r}
library(zstdlite)
library(lz4lite)
set.seed(1)

N                 <- 1e6
input_ints        <- sample(1:5, N, prob = (1:5)^4, replace = TRUE)
uncompressed      <- serialize(input_ints, NULL, xdr = FALSE)
compressed_lo     <- zstd_serialize(input_ints, level = -5)
compressed_mid    <- zstd_serialize(input_ints, level =  3)
compressed_mid2   <- zstd_serialize(input_ints, level = 10)
compressed_hi     <- zstd_serialize(input_ints, level = 22)
compressed_base   <- memCompress(serialize(input_ints, NULL, xdr = FALSE))
```


```{r echo = FALSE}
lens <- c(
  length(serialize(input_ints, NULL, xdr = FALSE)),
  length(memCompress(serialize(input_ints, NULL, xdr = FALSE))),
  length(zstd_serialize(input_ints, level = -5)),
  length(zstd_serialize(input_ints, level =  3)),
  length(zstd_serialize(input_ints, level = 10)),
  length(zstd_serialize(input_ints, level = 22))
)
```



<details>
<summary> Click here to show/hide benchmark code </summary>

```{r}
library(zstdlite)

res <- bench::mark(
  serialize(input_ints, NULL, xdr = FALSE),
  memCompress(serialize(input_ints, NULL, xdr = FALSE)),
  zstd_serialize(input_ints, level =  -5),
  zstd_serialize(input_ints, level =   3),
  zstd_serialize(input_ints, level =  10),
  zstd_serialize(input_ints, level =  22),
  check = FALSE
)
```
</details>


```{r echo = FALSE}
res %>% 
  mutate(`MB/s` = round(N*4/1024^2 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  mutate(compression_ratio = round(lens/(N*4), 3)) %>%
  select(expression, median, `itr/sec`, `MB/s`, compression_ratio) %>%
  knitr::kable()
```


### Decompressing 1 million integers


<details>
<summary> Click here to show/hide benchmark code </summary>

```{r}
res <- bench::mark(
  unserialize(uncompressed),
  zstd_unserialize(compressed_lo),
  zstd_unserialize(compressed_mid2),
  zstd_unserialize(compressed_hi),
  unserialize(memDecompress(compressed_base, type = 'gzip')),
  check = TRUE
)
```

</details>

```{r echo = FALSE}
res %>% 
  mutate(`MB/s` = round(N*4/1024^2 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  select(expression, median, `itr/sec`, `MB/s`)  %>%
  knitr::kable()
```



### Zstd "Single File" Libary

* To simplify the code within this package, it uses the 'single file library' version of zstd
* To update this package when zstd is updated, create the single file library version
    1. cd zstd/build/single_file_libs
    2. sh create_single_file_library.sh
    3. Wait.....
    4. copy zstd/built/single_file_libs/zstd.c into zstdlite/src
    5. copy zstd/lib/zstd.h into zstdlite/src

## Related Software


For a more general solution to fast serialization of R objects, see the 
[fst](https://github.com/fstpackage/fst) or [qs](https://cran.r-project.org/package=qs) packages.

* [lz4](https://github.com/lz4/lz4) and [zstd](https://github.com/facebook/zstd) - both by Yann Collet
* [fst](https://github.com/fstpackage/fst) for serialisation of data.frames using
  lz4 and zstd
* [qs](https://cran.r-project.org/package=qs) for fast serialization of arbitrary R objects
  with lz4 and zstd

## Acknowledgements

* Yann Collett for releasing, maintaining and advancing 
[lz4](https://github.com/lz4/lz4) and [zstd](https://github.com/facebook/zstd)
* R Core for developing and maintaining such a wonderful language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
